\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}

\newenvironment{ttdb_mmpdb}{\begin{verbatim}}{\end{verbatim}}
\newenvironment{ttdb_cursor}{\begin{verbatim}}{\end{verbatim}}
\newenvironment{ttdb_record}{\par\bigskip\hrule\medskip}{\medskip\hrule\bigskip}
\newenvironment{ttdb_special}{\begin{verbatim}}{\end{verbatim}}
\newcommand{\ttdbheader}[1]{\noindent\texttt{#1}\par\medskip}

\begin{document}
\section*{Mathematical TTDB}
This LaTeX TTDB specifies generalized mathematical processing for TTDB features used in \texttt{index.html} and \texttt{banjo/banjo.html}. It is intended as an implementation-neutral reference model.

\begin{ttdb_mmpdb}
db_id: ttdb:mathematical:processing:v1
db_name: "Mathematical"
coord_increment:
  lat: 1
  lon: 1
collision_policy: southeast_step
timestamp_kind: unix_utc
umwelt:
  umwelt_id: umwelt:tte:mathematical:model:v1
  role: process_librarian
  perspective: "A formal model of TTDB parsing, cursoring, rendering, interaction, and stochastic reveal."
  scope: "Features represented in index.html and banjo/banjo.html."
  constraints:
    - "Use deterministic transforms where possible."
    - "Preserve typed-edge directionality."
    - "Treat random card choice as explicit stochastic sampling."
  globe:
    frame: "unit_sphere_knowledge_globe"
    origin: "Record coordinates are points on a subjective globe."
    mapping: "Record ID @LATxLONy maps to spherical coordinates (phi, lambda) in degrees."
    note: "Projection and visibility are camera-dependent."
cursor_policy:
  max_preview_chars: 280
  max_nodes: 80
typed_edges:
  enabled: true
  syntax: "type>@TARGET_ID"
  note: "Edges are directed and canonical in header metadata."
librarian:
  enabled: true
  primitive_queries:
    - "SELECT <record_id>"
    - "FIND <token>"
    - "EDGES <record_id>"
    - "STATUS"
  max_reply_chars: 280
  invocation_prefix: "@AI"
\end{ttdb_mmpdb}

\begin{ttdb_cursor}
selected:
  - @LAT65LON45
preview:
  @LAT65LON45: "TTCP RFC series: normative requirements for rendering a TTDB file as an interactive web publication."
agent_note: "When ambiguity exists, prefer deterministic tie-break rules or explicit clarification."
\end{ttdb_cursor}

\begin{ttdb_record}
\ttdbheader{@LAT-90LON0 | created:1771899990 | updated:1771899990}
\begin{ttdb_special}
kind: discovery_tour_off
\end{ttdb_special}
\end{ttdb_record}

\begin{ttdb_record}
\ttdbheader{@LAT0LON0 | created:1771900000 | updated:1771901000 | relates:defines>@LAT18LON-120,defines>@LAT35LON-60,defines>@LAT52LON20,defines>@LAT28LON95,defines>@LAT5LON150,defines>@LAT-22LON60,defines>@LAT-42LON-30,defines>@LAT-12LON-150,defines>@LAT-62LON120,defines>@LAT65LON45}
\subsection*{System State and Dynamics}
Let the application state at time $t$ be
\[
S_t = (M_t, C_t, R_t, E_t, G_t, V_t, U_t),
\]
where:
\[
\begin{aligned}
M_t &: \text{mmpdb metadata}, \\
C_t &: \text{cursor state}, \\
R_t &: \text{record map}, \\
E_t &: \text{typed directed edges}, \\
G_t &: \text{record coordinate map}, \\
V_t &: \text{view/camera state}, \\
U_t &: \text{UI and timers}.
\end{aligned}
\]
An implementation evolves by
\[
S_{t+1} = F(S_t, a_t, \eta_t),
\]
with input action $a_t$ and optional stochastic term $\eta_t$ (used for random card draws).

\medskip\noindent\textbf{Normative references:} \texttt{TTCP-RFC-0001} (full pipeline); \texttt{TTCP-RFC-0002} (globe and navigation); \texttt{TTCP-RFC-0003} (link system and search).
\end{ttdb_record}

\begin{ttdb_record}
\ttdbheader{@LAT18LON-120 | created:1771900010 | updated:1771900010 | relates:refines>@LAT0LON0,feeds>@LAT35LON-60,feeds>@LAT52LON20}
\subsection*{Parsing and Canonicalization}
Define parser
\[
P : \Sigma^* \to (M, C, \mathcal{R}),
\]
where $\Sigma^*$ is file text and $\mathcal{R} = (r_1,\dots,r_n)$ is ordered records.
Each record is
\[
r_i = (id_i, h_i, b_i, title_i, X_i, \mathcal{E}_i).
\]
The ID grammar is
\[
id_i = \texttt{@LAT}\phi_i\texttt{LON}\lambda_i,\quad \phi_i,\lambda_i \in \mathbb{R}.
\]
Coordinate extraction is deterministic:
\[
X_i = (\phi_i, \lambda_i).
\]
For header token list $\texttt{relates:}$,
\[
\mathcal{E}_i = \{(\tau, id_j)\},
\]
with token syntax $\tau\texttt{>}id_j$ and default $\tau=\texttt{relates}$ if omitted.

\medskip\noindent\textbf{Normative references:} \texttt{TTCP-RFC-0001} Sections 2--4, 9 (file ingestion, three-zone parse, record header, LaTeX container).
\end{ttdb_record}

\begin{ttdb_record}
\ttdbheader{@LAT35LON-60 | created:1771900020 | updated:1771900020 | relates:refines>@LAT0LON0,depends_on>@LAT18LON-120,feeds>@LAT28LON95}
\subsection*{Typed-Edge Graph Semantics}
Define graph
\[
\mathcal{G}_{TTDB} = (\mathcal{V}, \mathcal{E}_{dir}),
\]
where $\mathcal{V}=\{id_i\}$ and
\[
\mathcal{E}_{dir} = \{(id_i,\tau,id_j)\mid (\tau,id_j)\in \mathcal{E}_i\}.
\]
Directionality rule:
\[
(id_i,\tau,id_j) \not\Rightarrow (id_j,\tau,id_i).
\]
Duplicate edges may be canonicalized by set projection:
\[
\mathcal{E}_{canon} = \text{unique}(\mathcal{E}_{dir}).
\]
Rendering may use $\mathcal{E}_{canon}$, but stored edge headers remain authoritative.

\medskip\noindent\textbf{Normative references:} \texttt{TTCP-RFC-0001} Section 12 (typed edge display); \texttt{TTCP-RFC-0002} Section 4 (edge rendering on globe).
\end{ttdb_record}

\begin{ttdb_record}
\ttdbheader{@LAT52LON20 | created:1771900030 | updated:1771900030 | relates:refines>@LAT0LON0,depends_on>@LAT18LON-120}
\subsection*{Cursor, Discovery, and Search}
Let ordered record IDs be $\mathcal{O}=(id_1,\dots,id_n)$ and selected ID be $s_t$.
Discovery set $D_t\subseteq \mathcal{O}$ updates by:
\[
D_{t+1} =
\begin{cases}
\mathcal{O}, & \text{if discovery mode is disabled},\\
D_t \cup \{id_1,s_t\}, & \text{otherwise}.
\end{cases}
\]
Search term $q$ yields visible order:
\[
\mathcal{F}(q)=
\begin{cases}
\mathcal{O}\cap D_t, & q=\varnothing,\\
\{id\in D_t\mid q\subseteq \text{blob}(id)\}, & q\neq\varnothing.
\end{cases}
\]
If $s_t\notin \mathcal{F}(q)$ and $\mathcal{F}(q)\neq\varnothing$, select first element.

\medskip\noindent\textbf{Normative references:} \texttt{TTCP-RFC-0002} Sections 6--7 (cursor selection, discovery); \texttt{TTCP-RFC-0003} Section 6 (search index and filtering).
\end{ttdb_record}

\begin{ttdb_record}
\ttdbheader{@LAT28LON95 | created:1771900040 | updated:1771900040 | relates:refines>@LAT0LON0,depends_on>@LAT35LON-60,feeds>@LAT5LON150}
\subsection*{Spherical Projection and Visibility}
Given $(\phi,\lambda)$ in degrees, convert to radians and map to the unit sphere:
\[
x=\cos\phi\sin\lambda,\quad y=\sin\phi,\quad z=\cos\phi\cos\lambda.
\]
Apply yaw $\rho$ then pitch $\theta$:
\[
\begin{bmatrix}x_1\\y_1\\z_1\end{bmatrix}
= R_x(\theta)R_y(\rho)\begin{bmatrix}x\\y\\z\end{bmatrix}.
\]
Node is front-visible iff $z_1>0$.
Screen projection for globe center $(c_x,c_y)$ and radius $r$:
\[
p_x=c_x+r x_1,\quad p_y=c_y-r y_1.
\]
This model is shared by TTDB globe and banjo globe rendering.

\medskip\noindent\textbf{Normative references:} \texttt{TTCP-RFC-0002} Sections 2--3 (globe projection model, node rendering, front-face culling).
\end{ttdb_record}

\begin{ttdb_record}
\ttdbheader{@LAT5LON150 | created:1771900050 | updated:1771900050 | relates:refines>@LAT0LON0,depends_on>@LAT28LON95,selects>@LAT52LON20}
\subsection*{Interaction Operators}
Nearest-node selection at pointer $p$:
\[
\hat{id}=\arg\min_{id\in \mathcal{V}_{screen}} \|s(id)-p\|_2,
\]
accepted when $\|s(\hat{id})-p\|_2\le \tau_{hit}$.

Drag updates camera:
\[
\rho \leftarrow \rho + k_x \Delta x,\quad
\theta \leftarrow \text{clamp}(\theta + \sigma k_y \Delta y,\;[-\tfrac{\pi}{2}+\epsilon,\tfrac{\pi}{2}-\epsilon]),
\]
where $\sigma\in\{-1,1\}$ encodes Y-axis inversion preference.

Zoom update:
\[
z \leftarrow \text{clamp}(z\cdot \gamma,\;[z_{min},z_{max}]).
\]

\medskip\noindent\textbf{Normative references:} \texttt{TTCP-RFC-0002} Section 5 (drag, zoom, tap, reduced motion).
\end{ttdb_record}

\begin{ttdb_record}
\ttdbheader{@LAT-22LON60 | created:1771900060 | updated:1771900060 | relates:refines>@LAT0LON0,depends_on>@LAT28LON95,depends_on>@LAT5LON150}
\subsection*{Record-to-Record Transition Dynamics}
For transition $id_a\to id_b$, direction uses longitude wrap delta $\Delta\lambda$ and latitude delta $\Delta\phi$:
\[
\mathbf{d} = (\Delta\lambda,-\Delta\phi),\quad \hat{\mathbf{d}}=\frac{\mathbf{d}}{\|\mathbf{d}\|_2}.
\]
Great-circle fraction
\[
\delta = \frac{2\arctan2\left(\sqrt{\alpha},\sqrt{1-\alpha}\right)}{\pi},
\]
\[
\alpha=\sin^2\!\left(\frac{\Delta\phi_r}{2}\right)+\cos\phi_{a,r}\cos\phi_{b,r}\sin^2\!\left(\frac{\Delta\lambda_r}{2}\right),
\]
clamped to $[0,1]$, then
\[
\delta'=\text{clamp}(\delta,[0.03,1]).
\]
Duration and travel:
\[
T=\text{round}\!\left((160+1400(\delta')^{0.55})\cdot \kappa\right),\quad
L=\text{round}\!\left(72+300(\delta')^{0.75}\right),
\]
where $\kappa$ is a pace multiplier (e.g., slow tour mode).

\medskip\noindent\textbf{Normative references:} \texttt{TTCP-RFC-0002} Sections 9.1--9.2 (transition parameters, named styles: slide, bloom, twist).
\end{ttdb_record}

\begin{ttdb_record}
\ttdbheader{@LAT-42LON-30 | created:1771900070 | updated:1771900070 | relates:refines>@LAT0LON0,depends_on>@LAT52LON20,depends_on>@LAT-22LON60}
\subsection*{Guided Tour Automaton}
Tour state machine:
\[
Q=\{\text{off},\text{paused},\text{scheduled}\}.
\]
If enabled and $|D_t|\ge 2$, schedule timer $\Delta t_{tour}$.
On timeout:
\[
next(s_t)=\mathcal{D}\left[(\text{index}(s_t)+1)\bmod |\mathcal{D}|\right],
\]
where $\mathcal{D}$ is discovered order.
Interaction events reset schedule (unless tour is off).

\medskip\noindent\textbf{Normative references:} \texttt{TTCP-RFC-0002} Section 8 (guided tour, tour audio, slow pace, space-key pause, discovery\_tour\_off suppression).
\end{ttdb_record}

\begin{ttdb_record}
\ttdbheader{@LAT-12LON-150 | created:1771900080 | updated:1771900080 | relates:refines>@LAT0LON0,depends_on>@LAT28LON95}
\subsection*{Banjo Three-Card Stochastic Reveal}
Let available set $A_t$ and drawn sequence $B_t$.
For each reveal step while $|B_t|<3$:
\[
c_t \sim \text{Uniform}(A_t).
\]
Rolling phase updates orientation:
\[
\rho \leftarrow \rho + \omega_\rho \Delta t,\quad
\theta \leftarrow \text{clamp}(\theta + \omega_\theta \Delta t + \epsilon\sin(\kappa t)\Delta t).
\]
At settle trigger, target orientation $(\theta^\star,\rho^\star)$ from card coordinate is approached by:
\[
\theta \leftarrow \theta + \beta\cdot \Delta\theta,\quad
\rho \leftarrow \rho + \beta\cdot \Delta\rho.
\]
When $|\Delta\theta|,|\Delta\rho|<\varepsilon$, reveal card:
\[
B_{t+1}=B_t \Vert [c_t],\quad A_{t+1}=A_t\setminus\{c_t\}.
\]

\medskip\noindent\textbf{Normative references:} \texttt{TTCP-RFC-0002} Section 2 (spherical projection model applies to banjo globe identically).
\end{ttdb_record}

\begin{ttdb_record}
\ttdbheader{@LAT-62LON120 | created:1771900090 | updated:1771900090 | relates:implements>@LAT18LON-120,implements>@LAT35LON-60,implements>@LAT52LON20}
\subsection*{Strategies for Rendering .latex TTDB in index.html}
\textbf{Strategy A: Build-time canonicalization.}
Convert \texttt{.latex} TTDB into canonical JSON:
\[
\texttt{latex} \xrightarrow{P_{latex}} \texttt{TTDB-JSON},
\]
then reuse existing renderer unchanged.

\textbf{Strategy B: Dual-parser client path.}
Detect extension and parse either markdown or LaTeX into the same in-memory model:
\[
\texttt{parseAny}(file)=
\begin{cases}
P_{md}(file), & \text{if .md}\\
P_{latex}(file), & \text{if .latex}
\end{cases}
\]
followed by shared rendering pipeline.

\textbf{Strategy C: On-load transpile to markdown subset.}
Map deterministic LaTeX environments to markdown-equivalent text and pass through current \texttt{parseRecords} with minimal UI changes.

\medskip\noindent\textbf{Normative references:} \texttt{TTCP-RFC-0001} Sections 2, 9 (file ingestion and format detection; LaTeX container support and body conversion rules).
\end{ttdb_record}

\begin{ttdb_record}
\ttdbheader{@LAT65LON45 | created:1771901000 | updated:1771901000 | relates:specified_by>@LAT0LON0,covers>@LAT18LON-120,covers>@LAT35LON-60,covers>@LAT52LON20,covers>@LAT28LON95,covers>@LAT5LON150,covers>@LAT-22LON60,covers>@LAT-42LON-30,covers>@LAT-62LON120}
\subsection*{Content Publishing RFC Series (TTCP)}
The TTCP RFC series formalizes the normative requirements for rendering a TTDB file as an interactive web publication. Three documents constitute the series:

\medskip
\noindent\texttt{TTCP-RFC-0001} --- \textbf{Record Rendering.}
File ingestion (fetch, format detection); three-zone parse model (mmpdb, cursor, records); record header fields; Markdown and LaTeX rendering pipelines; media dispatch and lead-media extraction; split layout; \texttt{ttdb-record} audio config; special records (\texttt{tour\_sound}, \texttt{discovery\_tour\_off}); epistemic weight display; typed edge display.

\medskip
\noindent\texttt{TTCP-RFC-0002} --- \textbf{Globe and Navigation.}
Spherical projection model; node states (undiscovered, discovered, selected eyeball); deterministic discovery colors; edge rendering; drag, zoom, and tap interaction; cursor selection and globe rotation animation; discovery system and localStorage persistence; guided tour automaton; record transition animation (slide, bloom, twist); scene record playback (\texttt{ttdb-scene}); multi-database side-globe navigation.

\medskip
\noindent\texttt{TTCP-RFC-0003} --- \textbf{Link System and Addressability.}
\texttt{toot:} URI scheme and record token format; four supported link-target forms (toot URI, legacy \texttt{@}, path, viewer URL); DB alias resolution; URL synchronization via \texttt{history.replaceState}; search index construction and substring filtering over discovered records.

\medskip
\noindent The mapping from state-space components defined in \texttt{@LAT0LON0} to TTCP RFC sections is:
\[
\begin{aligned}
M_t, C_t, R_t &\mapsto \texttt{TTCP-RFC-0001} \text{ (parse and render)}, \\
E_t           &\mapsto \texttt{TTCP-RFC-0001 \S12},\ \texttt{TTCP-RFC-0002 \S4}, \\
G_t, V_t      &\mapsto \texttt{TTCP-RFC-0002} \text{ (globe and navigation)}, \\
U_t           &\mapsto \texttt{TTCP-RFC-0002 \S5--11},\ \texttt{TTCP-RFC-0003}.
\end{aligned}
\]
\end{ttdb_record}

\end{document}
